---
title: "Reproducible Research: Peer Assessment 1"
author: "Eric Guo"
date: "Monday, December 08, 2014"
output: html_document
keep_md: true
---

```{r echo=FALSE, results="hide", message=FALSE}
library(dplyr);
library(ggplot2);
```

## Loading and preprocessing the data.

```{r cache=TRUE}
unzip("activity.zip");
datas <- read.csv(file = "activity.csv", header = TRUE);
```

## What is mean total number of steps taken per day?

```{r}
data <- datas;
data <- group_by(data, date);
data <- summarise(data, count = sum(steps, na.rm = TRUE));

hist(data$count, 
     main = "Histogram of the total number of steps taken each day",
     xlab = "Total number of steps taken each day");
```

The mean of total number of steps taken per day is `r format(round(mean(data$count), 2), nsmall = 2)`. The median of total number of steps taken per day is `r median(data$count)`.

## What is the average daily activity pattern?

```{r}
data <- datas;
data <- group_by(data, interval);
data <- summarise(data, mean = mean(steps, na.rm = TRUE));

ggp <- ggplot(data, aes(x = interval, y = mean)) + 
        geom_line(stat = "identity") +
        labs(x = "Interval", y = "Number of steps") +
        labs(title = "The 5-minute interval and the average number of steps taken");
print(ggp);
```

The 5-minute interval '`r data[data$mean == max(data$mean), 1]`' , on average across all the days in the dataset, contains the maximum number of steps.

## Imputing missing values.

(1). The total number of missing values in the dataset.

```{r}
data <- datas;
data <- data[is.na(data$steps),];
print(length(data$steps));

```

(2). Filling in all of the missing values in the dataset. 

```{r}
data <- datas;
data[is.na(data$steps), c("steps")] <- as.integer(mean(datas$steps, na.rm = TRUE));

```

(3). Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
data_new <- data;

```

(4). Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

```{r}
data <- data_new;
data <- group_by(data, date);
data <- summarise(data, count = sum(steps, na.rm = TRUE));

hist(data$count, 
     main = "Histogram of the total number of steps taken each day",
     xlab = "Total number of steps taken each day");
```

The mean of total number of steps taken per day is `r format(round(mean(data$count), 2), nsmall = 2)`. The median of total number of steps taken per day is `r median(data$count)`. These values are differ from the estimates from the first part of the assignment, and they are increased.

## Are there differences in activity patterns between weekdays and weekends?

(1). Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r results="hide"}
Sys.setlocale(category = "LC_TIME", locale = "English");
data <- data_new;
data <- mutate(data, weekday = weekdays(as.Date(date)));
data[grepl("Saturday", data$weekday) | grepl("Sunday", data$weekday), c("weekday")] <- "weekend";
data[!grepl("weekend", data$weekday), c("weekday")] <- "weekday";
data$weekday <- as.factor(data$weekday);
```

(2). Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r}
data <- group_by(data, weekday, interval);
data <- summarise(data, mean = mean(steps, na.rm = TRUE));

ggp <- ggplot(data, aes(x = interval, y = mean)) + 
        geom_line(stat = "identity") +
        facet_grid(weekday~., space = "free") +
        labs(x = "Interval", y = "Number of steps") +
        labs(title = "The 5-minute interval and the average number of steps taken"); 
print(ggp);
```


